The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2010-12-03 smurf_paper.tex
add definition of \prd for Stompor et al.
describe PCA analysis

% journals
\newcommand{\apj}{\rm ApJ}
\newcommand{\apjl}{\rm ApJL}
\newcommand{\apjs}{\rm ApJS}
\newcommand{\aaps}{\rm A$\&$AS}
\newcommand{\aap}{\rm A$\&$A}
\newcommand{\aapr}{\rm A$\&$AR}
\newcommand{\mnras}{\rm MNRAS}
\newcommand{\aj}{\rm Astron. J.}
\newcommand{\araa}{\rm ARAA}
\newcommand{\nat}{\rm Nature}
\newcommand{\pasj}{\rm PASJ}
\newcommand{\pasp}{\rm Publ. Astron. Soc. Pac.}
\newcommand{\ASP}{\rm ASP COnference Series}
\newcommand{\CASP}{\rm Comm. Astrophys. Space Phys.}
\newcommand{\astroph}{\rm astro-ph/}

@@

The scan strategy is designed, primarily, to mitigate the effects of
$1/f$ drift. The simplest scan one might use is a raster pattern, or a
series of parallel passes across this field. However, such a scan
yields extremely poor performace for structures that are transverse to
the scan direction, since a long time will have ellapsed between each
parallel step, and the signal will therefore be totally dominated by
the $1/f$ noise. In addition, other noise features may appear in the
data periodically. With such a regularly spaced periodic scan
strategy, these noise features can easily conspire with the scan
pattern to produce similarly periodic noise structures in the map, or
\emph{scan-synchronous} noise. It has therefore been recognized for
some time that scan strategies that visit each region of the map on as
many different time scales, and scan angles as possible offer the best
protection against large-scale noise (...cite some people...). As a
tradeoff between truly random scan patterns and feasibility of
implementation taking into consideration both mechanical and
scheduling constraints, SCUBA-2 presently has three primary scan
modes: PONG (cite Borys et al.), Lissajous, and Daisy... (more
citations? Kovacs et al. has a nice description... also CMB
literature).

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pca.pdf}
\caption{The first six modes from a principal component analysis of
  the combined 450 and 850\,\micron\ time-series bolometer data.}
\label{fig:pca}
\end{figure*}

\subsubsection{Sky noise}
Do we even see sky noise? Can show some results for the combined 450
and 850\,\micron\ joint PCA analysis.

@@

\subsubsection{Line features}

The power spectra also show some narrower line features (many
correlated across the array). Even though we largely ignore them worth
some mention.

>>>>

% journals
\newcommand{\apj}{\rm ApJ}
\newcommand{\apjl}{\rm ApJL}
\newcommand{\apjs}{\rm ApJS}
\newcommand{\aaps}{\rm A$\&$AS}
\newcommand{\aap}{\rm A$\&$A}
\newcommand{\aapr}{\rm A$\&$AR}
\newcommand{\mnras}{\rm MNRAS}
\newcommand{\aj}{\rm Astron. J.}
\newcommand{\araa}{\rm ARAA}
\newcommand{\nat}{\rm Nature}
\newcommand{\pasj}{\rm PASJ}
\newcommand{\pasp}{\rm Publ. Astron. Soc. Pac.}
\newcommand{\prd}{\rm PhRvD}
\newcommand{\ASP}{\rm ASP Conference Series}
\newcommand{\CASP}{\rm Comm. Astrophys. Space Phys.}
\newcommand{\astroph}{\rm astro-ph/}

@@

The scan strategy is designed, primarily, to mitigate the effects of
$1/f$ drift. The simplest scan one might use is a raster pattern, or a
series of parallel passes across this field. However, such a scan
yields extremely poor performace for structures that are transverse to
the scan direction, since a long time will have ellapsed between each
parallel step, and the signal will therefore be totally dominated by
the $1/f$ noise. In addition, other noise features may appear in the
data periodically. With such a regularly spaced periodic scan
strategy, these noise features can easily conspire with the scan
pattern to produce similarly periodic noise structures in the map, or
\emph{scan-synchronous} noise. It has therefore been recognized for
some time that scan strategies that visit each region of the map on as
many different time scales, and scan angles as possible offer the best
protection against large-scale noise (...cite some people...); such
strategies are said to offer good \emph{cross-linking}. As a tradeoff
between truly random scan patterns and feasibility of implementation
taking into consideration both mechanical and scheduling constraints,
SCUBA-2 presently has three primary scan modes: PONG (cite Borys et
al.), Lissajous, and Daisy... (more citations? Kovacs et al. has a
nice description... also CMB literature).

@@

A method that has been employed by teams analyzing BOLOCAM and AzTEC
data to identify and remove low-frequency correlated noise is
Principal Component Analysis
\citep[PCA,][]{laurent2005,scott2008}. While we have not (yet)
included ``PCA cleaning'' as a standard feature of SMURF, we have
performed a preliminary analysis in IDL of some SCUBA-2 data to gain
further insight into the nature of the low-frequency noise. The basic
method is as follows: (i) the covariance matrix is built up for all
pairs $(i,j)$ of the $N$ bolometer time series,
$\langle\mathbf{b}_i(t),\mathbf{b}_j\rangle$; and (ii) an eigenvalue
decomposition is used to identify a new set of statistically
uncorrelated eigenvectors, $\mathbf{\xi}_i(t)$ (i.e. whose covariance
matrix is diagonal), such that each of the bolometer time series is a
linear combination of the eigenvectors, or \emph{components}
i.e. $\mathbf{b}_i(t) = \bar{\mathbf{\xi}} \mathbf{\lambda}_i^T$,
where each row of the matrix $\bar{\mathbf{\xi}}$ is an eigenvector,
and $\mathbf{\lambda}_i^T$ is a column vector containing the
corresponding eigenvalues. For convenience, the $\mathbf{\xi}_i$ are
normalized by their \rms, so that the amplitude of each component is
stored in the eigenvalues. In the earlier analyses mentioned, the
low-frequency noise is assumed to be encapsulated in those components
with the largest eigenvalues. Removing the projection of the time
series along those components has been used successfully to reduce
$1/f$ noise while retaining most of the signal in point-like sources.

One unique feature of the SCUBA-2 data is that one can perform a PCA
analysis of both the 450 and 850\,\micron\ data simultaneously. The
first 6 most significant modes of our analysis are shown in
Fig.~\ref{fig:pca}, with the eigenvectors in the top panels, and maps
of the eigenvalues in each subarray in the bottom panels.

The majority of correlated signal at both wavelengths is clearly
broken down into the first two plotted components. The first dominates
at 450\,micron, whereas the second dominates at 850\,micron. For
reference, the eigenvalues of these two modes are in the range
$\sim1$--4\,pW, whereas the remaining modes have eigenvalues
$\lsim0.3$\,pW. As mentioned earlier, one simple model for the
correlated signal is a simple combination of oscillations in base
temperature (which would be similar in each band), with a varying
atmospheric signal (which is expected to be stronger at
450\,\micron). Indeed, the fainter component 2 most closely resembles
the mixing chamber temperature variations (Fig.~\ref{fig:bolos_mix}),
whereas component 1 is stronger at 450\,\micron\ and does not resemble
the fridge signal as closely.

This physical interpretation of the eigenvectors should, however, be
taken with a grain of salt, since the PCA analysis is really a
statistical black box. Suppose most of the low frequency noise really
could be decomposed into two components, the fridge signal
$\mathbf{F}(t)$, and an atmospheric signal $\mathbf{A}(t)$. The two
eigenvectors obtained from the PCA analysis are simply linear
combinations of $\mathbf{F}(t)$ and $\mathbf{A}(t)$ that are
statistically uncorrelated. Depending on the particular realizations
of these two noise sources, this decomposition may, or may not give
rise to eigenvectors that resemble the two underlying physical
sources. Regardless, similar analyses of a few data sets during S2SRO
give comparable results; most of the signal at both 450\,\micron\ and
850\,\micron\ appears to be produced by two dominant underlying
mechanisms.

@@
